On the Alfvenicity of solar wind discontinuities

### 2025-01-12 article.qmd
docs: update

- Observation features of solar wind discontinuities
- Importance of solar wind discontinuities
- The necessity of kinetic models to describe solar wind discontinuities
- Current theoretical models of solar wind discontinuities and their limitations
  - Harris current sheet @harrisPlasmaSheathSeparating1962

@@

Ampere’s law connects the fields and the flow components

$$
d B_y / d z = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n u_x + Γ_e B_x / B_z
$$ {#eq-Jx}

$$
d B_x / d z = J_y = \sum_α n_α u_{α y} - n_e u_{e y} = n u_y - Γ_e B_y / B_z
$$ {#eq-Jy}

$$
0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n u_z - Γ_e
$$ {#eq-Jz}

@@

$$
- u_α \sin θ_α θ_α' = n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ
$$

$$
\Rightarrow 
θ_α' = - \frac{n_α B_z}{Γ_α} + \frac{B_0}{u_α}
$${#eq-theta1}

Note that $θ_α, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system. So given the profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ_α$, thus the profile of $u_α$.

@@

$$
- B_0 \sin θ θ' = \sum n_α u_{α} \sin θ_α - Γ_e B_0 \sin θ / B_z
$$

$$
\Rightarrow 
θ' = - \frac{ \sum n_α u_{α} }{B_0} + \frac{Γ_e}{B_z}
$${#eq-theta}

@@

$$
n u = B_0 \left(\frac{Γ_e}{B_z}+\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)
$$

In the asymptotic region, as all variables approach constant values, the derivatives must vanish. For @eq-theta1, this means $θ_α(∞)' = 0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$. This relates the velocity of each species to the asymptotic density. Rewriting the @eq-theta1 in terms of asymptotic values, we have:

@@

$$
\begin{aligned}
B_x(z) &= B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
B_y(z) &= B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
n(z) &= n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
\\
J_x(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
J_y(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$

>>>>

<!-- - Observation features of solar wind discontinuities
- Importance of solar wind discontinuities
- The necessity of kinetic models to describe solar wind discontinuities
- Current theoretical models of solar wind discontinuities and their limitations
  - Harris current sheet @harrisPlasmaSheathSeparating1962 -->

@@

Ampere's law connects the fields and the flow components

$$
\frac{1}{μ_0} \frac{d B_y}{d z}  = - J_x = - \sum_α n_α u_{α x} + n_e u_{e x} = -n U_x + Γ_e B_x / B_z
$$ {#eq-Jx}

$$
\frac{1}{μ_0} \frac{d B_x}{d z} = J_y = \sum_α q_α n_α u_{α y} - e n_e u_{e y} = e (n U_y - Γ_e \frac{B_y}{B_z})
$$ {#eq-Jy}

$$
0 = J_z = \sum_α n_α u_{α z} - n_e u_{e z} = n U_z - Γ_e
$$ {#eq-Jz}

@@

$$
- u_α \sin θ_α θ_α' = \frac{e}{m_α} (n_α u_α \sin θ_α B_z / Γ_α - B_0 \sin θ)
$$

$$
\Rightarrow 
θ' = \frac{e}{m_α} \left(\frac{B_0}{u_α} - \frac{n_α B_z}{Γ_α}\right)
$${#eq-theta1}
where $'=d/dz$.

Note that $θ, n_α$ are dependent variables, and $u_α, B_0, B_z, Γ_α$ are constants pre-determined by the system. So for a given profile of $n_α$ and the boundary condition, we could solve the above equation to get the profile of $θ$.

@@

$$
- B_0 \sin θ θ' = μ_0 e \left(\sum n_α u_{α} \sin θ_α - Γ_e  \sin θ \frac{B_0}{B_z}\right)
$$

$$
\Rightarrow 
θ' = μ_0 e \left(\frac{Γ_e}{B_z} - \frac{ \sum n_α u_{α} }{B_0}\right)
$${#eq-theta}

@@

$$
n U = B_0 \left(\frac{Γ_e}{B_z}+\frac{1}{μ_0 m_α}\left(\frac{n_α B_z}{Γ_α}-\frac{B_0}{u_α}\right)\right)
$$

In the asymptotic region, as all variables approach constant values, the derivatives must vanish. For @eq-theta1, this means $0 = - \frac{n_α(∞) B_z}{Γ_α} + \frac{B_0}{u_α}$. This relates the velocity of each species to the asymptotic density. Rewriting the @eq-theta1 in terms of asymptotic values, we have:

@@

$$
\begin{aligned}
B_x(z) &= B_0 \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
B_y(z) &= B_0 \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
n(z) &= n(\infty ) + \frac{\kappa ∑_α Γ_α}{1+(z/L)^2}
\\
J_x(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \sin \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\\
J_y(z) &= \frac{\kappa B_0 B_z}{1+(z/L)^2} \cos \left(\kappa  L B_z \tan ^{-1}\left(\frac{z}{L}\right)\right)
\end{aligned}
$$
where $\sum_α Γ_α=\sum_α n_α(∞)u_{z,α}(∞)$ and $i$ represents either $x$ or $y$. Using the relation that $Δn_α / Δn = Γ_α / \sum_α Γ_α$ and Equation~\ref{eq-asym1}, $\sum_α Γ_α$ can be reformulated into a form that does not involve $u_{z,α}(∞)$, which is often challenging to measure. This alternative expression is given by:

$$
\sum_α Γ_α = \sqrt{ \frac{B_z^2}{μ_0 m_p} \frac{{Δn}^2}{\sum_α {Δn_α}^2/ n_α(∞)} }
$$

For the simplest situation where we have two populations of ions of the same density but with oppositely directed bulk velocity, i.e., \(n_1 = n_2, u_1 = -u_2\), the system could be determined given $\frac{\sum_α Γ_α}{Δn}$, \(L\), and \(B_0\). The system profile of magnetic field, plasma density, and plasma velocity for one specific case ($L=d_i, \sum_α Γ_α= Δn V_A, B_0 = 2 B_z$) is plotted below Figure~\ref{fig-profiles}. The specific profiles are chosen to have zero electron current across the current sheet and zero \(B_y\) in the asymptotic limit (far away from the current sheet), corresponding to a $180^\circ$ rotation of the magnetic field across the current sheet. However, in general, we would expect the electron current to be non-zero.
